Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-22 paper.tex
fix typos
progress on writing

\newcommand{\typingHC}[3]{#1 : #2 \Longrightarrow #3}
\newcommand{\hcvOOinj}[2]{\mathtt{inj} \; #2}
%\newcommand{\hcvOOfun}[5]{\mathtt{fun} \; #1 \; #2 \; #3 \; #4 \; #5}
\newcommand{\hcvOOfun}[5]{\mathtt{fun} \; #2 \; #1 \; #3 \; #4 \; #5}
\newcommand{\hcvOOtt}[0]{\mathtt{unit}}
%\newcommand{\hcvOOcons}[4]{\mathtt{cons}\;#1\;#2\;#3\;#4}
\newcommand{\hcvOOcons}[4]{\mathtt{cons}\;#1\langle#2\rangle\;#3\langle#4\rangle}
%\newcommand{\hcvOOinl}[2]{\mathtt{inl}\;#1\;#2}
%\newcommand{\hcvOOinr}[2]{\mathtt{inr}\;#1\;#2}
\newcommand{\hcvOOinl}[2]{\mathtt{inl}\;#1\langle#2\rangle}
\newcommand{\hcvOOinr}[2]{\mathtt{inr}\;#1\langle#2\rangle}
\newcommand{\hckOOmt}[0]{\mathtt{stop}}
\newcommand{\hckOOconsI}[3]{\mathtt{cons_1}\;#1\;#2\;#3}
\newcommand{\hckOOappII}[2]{\mathtt{app_2}\;#1\;#2}
\newcommand{\sidecond}[1]{\text{if}\;#1}
% Lazy D cast calculus on space-inefficient CEK
\newcommand{\judgeCreduce}[2]{#1 \longmapsto_{\mathcal{C}} #2}
\newcommand{\judgeCreduceTrans}[2]{#1 \longmapsto_{\mathcal{C}}^{*} #2}
\newcommand{\judgeCeval}[2]{eval_{\mathcal{C}}(#1) = #2}
\newcommand{\redrule}[3]{#1 & \longmapsto_\mathcal{C} & #2 & #3\\}
\newcommand{\hiredrule}[3]{\highlight{#1} & \highlight{\longmapsto_\mathcal{C}} & \highlight{#2} & \highlight{#3} \\}
% blame calculus on space-efficient CEK
\newcommand{\judgeSreduce}[3]{#2 \longmapsto_{\mathcal{S}(#1)} #3}
\newcommand{\judgeSreduceTrans}[3]{#2 \longmapsto_{\mathcal{S}(#1)}^{*} #3}
\newcommand{\judgeSeval}[3]{eval_{\mathcal{S}(#1)}(#2) = #3}
\newcommand{\redruleS}[3]{#1 & \longmapsto_{\mathcal{S}(C)} & #2 & #3\\}
\newcommand{\hiredruleS}[3]{\highlight{#1} & 
\highlight{\longmapsto_{\mathcal{S}(C)}} & \highlight{#2} & \highlight{#3} \\}
% Normal Coercion
\newcommand{\ncProj}[2]{#1?^{#2}}
\newcommand{\ncInj}[1]{#1!}
\newcommand{\ncId}[0]{\iota}
\newcommand{\ncSeq}[2]{#1;#2}
\newcommand{\ncFail}[1]{\bot^{#1}}
\newcommand{\ncFun}[2]{\POOfun{#1}{#2}}
\newcommand{\ncProd}[2]{\POOprod{#1}{#2}}
\newcommand{\ncSum}[2]{\POOsum{#1}{#2}}
% Hypercoercion
\newcommand{\hyperCoercionI}[0]{\mathtt{id\star}}
\newcommand{\hyperCoercionC}[3]{#1 \overset{#2}{\curvearrowright} #3}
% machine state simulations
\newcommand{\eqvS}[4]{#3 \approx_{\mathcal{S}\mathcal{S}} #4}
\newcommand{\eqvSD}[3]{#2 \approx_{\mathcal{SD}} #3}
% to-dos
\newcommand{\todoKC}[1]{\todo[inline]{KC needs to #1}}
\newcommand{\todoKCFixed}[0]{\todo[inline]{Fixed. -KC}}
% abbrev
\newcommand{\castCalculus}[0]{$\lambda_{\rightarrow}^{\langle\cdot\rangle}$}
% names
\newcommand{\ineffCEK}{$\mathcal{C}$}
\newcommand{\ineffCEKD}{$\mathcal{D}$}
\newcommand{\ineffCEKUD}{$\mathcal{UD}$}
\newcommand{\judgeDreduce}[2]{#1 \longmapsto_{\mathcal{D}} #2}
\newcommand{\judgeDreduceTrans}[2]{#1 \longmapsto_{\mathcal{D}}^{*} #2}
\newcommand{\judgeDeval}[2]{eval_{\mathcal{D}}(#1) = #2}
\newcommand{\judgeUDreduce}[2]{#1 \longmapsto_{\mathcal{UD}} #2}
\newcommand{\judgeUDreduceTrans}[2]{#1 \longmapsto_{\mathcal{UD}}^{*} #2}
\newcommand{\judgeUDeval}[2]{eval_{\mathcal{UD}}(#1) = #2}
\newcommand{\effCEK}[1]{$\mathcal{S}(#1)$}
\newcommand{\evalEqv}[2]{$eval_{\text{#1}} = eval_{\text{#2}}$}
\newcommand{\continue}[2]{cont(#2,#1)}
\newcommand{\highlight}[1]{{\color{red} #1}}

@@

\begin{definition}[\lazyD{} CEK Machine]
  The \lazyD{} CEK machine, written \ineffCEKD{}, is the CEK machine of
  \figref{fig:machine-cekc} using the $applyCast$ for \lazyD{} defined
  in \figref{fig:applyCast-D-C}.  We write the transition relations of
  this machine as $\judgeUDreduce{s}{s}$ and $\judgeUDreduceTrans{s}{s}$
  and write the evaluation function as $\judgeUDeval{e}{o}$.
\end{definition}

@@

\begin{definition}[\lazyUD{} CEK Machine]
  The \lazyUD{} CEK machine, written \ineffCEKUD{}, is the CEK machine of
  \figref{fig:machine-cekc} using the $applyCast$ for \lazyUD{} defined
  in \figref{fig:apply-Cast-UD}.  We write the transition relations of
  this machine as $\judgeDreduce{s}{s}$ and $\judgeDreduceTrans{s}{s}$
  and write the evaluation function as $\judgeDeval{e}{o}$.
\end{definition}

@@

This section presents our second contribution, a framework for proving
correctness of cast representations, especially space efficient
ones. Section~\ref{sec:hypercoercion-correctness} will apply this
framework to prove the correctness of the \lazyD{} hypercoercions.

We start by defining a CEK machine \effCEK{C} that is parameterized
over the cast representation $C$. This machine is space efficient
provided that the cast representation performs compression.  The
interface between the machine and the cast representation is defined
by an abstract data type, named Cast ADT, that we define shortly.

We prove that, for any instance $C$ of the Cast ADT, \effCEK{C} is
equivelent to \ineffCEKD{] provided that $C$ satisfies a collection of
properties that we call the Lazy D Cast abstract data type.  In
Section~\ref{sec:hypercoercion-correctness} we show that the \lazyD{}
hypercoercions $H$ are an instance of the Lazy D ADT, and therefore
\effCEK{H} is equivalent to \ineffCEKD{}.

We are working on a similar theorem for \lazyUD{}.

We captures the set of operators that every cast representation must provide 
with \emph{cast abstract data type} (Definition~\ref{def:cast-rep}). 
The first three operators allow \effCEK{C} to construct casts. 
So we call them \textit{cast constructors}. And the forth and last operator 
allows \effCEK{C} to apply casts onto values.

\begin{definition}[Cast Abstract Data Type (Cast ADT)]
	\label{def:cast-rep}
	A cast abstract data type is a set $Cast$, which is indexed by two types, 
	with four operators:
	\begin{description}
		\item[$id(T)$] constructs an cast from a type
%		, where $c : T 
%		\Longrightarrow T $
		\item[$seq(c_1,c_2)$] composes two casts
%		, where $
%		\judgeTypeFT{c_1}{T_1}{T_2}$, $\judgeTypeFT{c_2}{T_2}{T_3}$, and $
%		\judgeTypeFT{c}{T_1}{T_3} $
		\item[$cast(T_1,l,T_2)$] constructs a cast from $T_1$ to $T_2$
		\item[$ applyCast(v,c)=r$] applies a cast onto a value
	\end{description}
	We denote by $c : T_1 \Longrightarrow T_2$ to mean $c$ is in $Cast \; 
	T_1 \; T_2$. And we say ``$c$ is from $T_1$ to $T_2$''.
\end{definition}

To use our framework, users needs to prove that their cast representation
implements Lazy D Cast ADT (Definition~\ref{def:surely-lazyd}), a subset of 
Cast ADT. Property (1) 
states that $id(T)$ acts like the identity function. Property (2) states 
that $seq(c,c)$ acts like a sequence of casts. Properties (3) through 
(11) state that $cast(T_1,l,T_2)$ act like the Lazy D $applyCast$
(Fig.~\ref{fig:applyCast-D-C}).

\todo[inline]{We might need to values before Lazy D Cast ADT. This will reduce 
the extreme height of Fig.~\ref{fig:machine-cekcc} as well. -KC}

\begin{definition}[Lazy D Cast ADT]
	\label{def:surely-lazyd}
	A  Cast is a Lazy D Cast if:
	\begin{enumerate}
		\item if $v : T$, then$applyCast(v,id(T)) = \mathtt{succ} \; v $
		\item if $\judgeType{v}{T_1}$,
		$ \judgeTypeFT{c_1}{T_1}{T_2}$, and
		$ \judgeTypeFT{c_2}{T_2}{T_3}$,\\
		then$applyCast(v,seq(c_1,c_2)) = 
		applyCast(v,c_1) >>= \lambda v.applyCast(v,c_2)$\\
		where 
		\[
		\begin{array}{rcl}
		\rOOsucc{v} >>= f & = & f(v) \\
		\rOOfail{l} >>= f & = & \rOOfail{l}
		\end{array}
		\]
		\item if $v : T_1$ and $\neg T_1 \smile T_2$,
		then$applyCast(v,cast(T_1, l, T_2)) = \rOOfail{l} $
		\item if $v : \star$, 
		then$applyCast(v,cast(\TOOdyn,l,\TOOdyn)) = \rOOsucc{v} $
		\item if $v : P_1$,
		then$applyCast(\hcvOOinj{P_1}{v},cast(\star,l,P_2)) 
		= applyCast(v,cast(P_1,l,P_2)) $
		\item if $v : P$,
		then$applyCast(v,cast(P,l,\star)) = \rOOsucc{(\hcvOOinj{P}{v})} $
		\item if $v : \POOunit$,
		then$applyCast(v,cast(\POOunit,l,\POOunit)) = \rOOsucc{v} $
		\item if $(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}) : \POOfun{T_1}{T_2}$,
		then\\
		$ 
		applyCast(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}, 
		cast(\POOfun{T_1}{T_2},l,\POOfun{T_3}{T_4})) \\
		= 
		\rOOsucc{(\hcvOOfun{seq(cast(T_3,l,T_1),c_1)}{\rho}{x}{b}{seq(c_2,cast(T_2,l,T_4))})}$
		\item if $(\hcvOOcons{v_1}{c_1}{v_2}{c_2}) : \POOprod{T_1}{T_2}$,
		then \\
		$ 
		applyCast(\hcvOOcons{v_1}{c_1}{v_2}{c_2},cast(\POOprod{T_1}{T_2},l,T_3 
		\times 
		T_4))$\\
		$ = 
		\rOOsucc{(\hcvOOcons{v_1}{seq(c_1,cast(T_1,l,T_3))}{v_2}{seq(c_2,cast(T_2,l,T_4))})}
		$ 
		\item if $(\hcvOOinl{v}{c}) : \POOsum{T_1}{T_2}$,
		then \\
		$ 
		applyCast(\hcvOOinl{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
		= \rOOsucc{(\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))})} $
		\item if $(\hcvOOinr{v}{c}) : \POOsum{T_1}{T_2}$,
		then \\$
		applyCast(\hcvOOinr{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
		= \rOOsucc{(\hcvOOinr{v}{seq(c,cast(T_2,l,T_4))})} $
	\end{enumerate}
\end{definition}

After showing their cast representation is a Lazy D Cast, users of our 
frameworks can plug-in this cast representation to our space-efficient abstract 
machine in subsection~\ref{ssec:framework:cek}, and prove that this machine is 
equivalent to \ineffCEKD\ (subsection~\ref{secc:framework:all-correct}), which 
depends on lemmas in subsection~\ref{secc:framework:monoid-correct}.

@@

\begin{figure}
	Syntax
	\[
	\begin{array}{rclr}

	\stxrule{v}{values}{
		\hcvOOtt \mid
		\hcvOOfun{c}{\rho}{x}{e}{c} \mid
		\hcvOOcons{v}{c}{v}{c}
	}
	\stxrulecont{
		\hcvOOinl{v}{c} \mid
		\hcvOOinr{v}{c} \mid
		\hcvOOinj{P}{v}
	}
%	\stxrule{r}{cast results}{
%		\rOOsucc{v} \mid
%		\rOOfail{l}
%	}
%	\stxrule{s}{states}{
%		\sOOinspect{e}{\rho}{\kappa} \mid{}
%		\sOOreturn{v}{\kappa} \mid{}
%		\sOOhalt{o}
%	}
	\stxrule{\kappa}{continuation}{
		\kOOcast{c}{k}
	}
	\stxrule{k}{pre-continuations}{
		\kOOmt \mid{}
		\kOOconsI{e}{\rho}{\kappa} \mid
		\kOOconsII{v}{\kappa} \mid
		\kOOinl{\kappa} \mid
		\kOOinr{\kappa}
	}
	\stxrulecont{
		\kOOappI{e}{\rho}{\kappa} \mid
		\kOOappII{v}{\kappa} \mid
		\kOOcar{\kappa} \mid
		\kOOcdr{\kappa}
	}
	\stxrulecont{
		\kOOcaseI{e}{e}{\rho}{\kappa} \mid
		\kOOcaseII{v}{e}{\rho}{\kappa} \mid
		\kOOcaseIII{v}{v}{\kappa}
	}
	\end{array}
	\]

@@

	\caption{Space-efficient CEK machine$\mathcal{S}(C)$}
	\label{fig:machine-cekcc}
\end{figure}
\figref{fig:machine-cekcc} defines \effCEK{C}, a space-efficient
CEK machine.  This machine is similar to \ineffCEK{}.
The keys differences are in values and continuations.

Let $v$ range over value. 
In \ineffCEK\, we have one value constructor for all casted value.
In \effCEK{C}, however, we do not have this generic value constructor, 
instead we push those casts into the ordinary values. For 
instance, $(\vOOcons{v_1}{v_2})$ corresponds to 
$(\hcvOOcons{v_1'}{id(T_1)}{v_2'}{id(T_2)})$. To construct values of the 
dynamic type, we use $(\hcvOOinj{P}{v})$.

Cast results ($r$) and machine states ($s$) are as before.

Let $\kappa$ range over continuations and Let $k$ range over 
pre-continuations. 
A continuation is now a pre-continuation prefixed with a cast.
Pre-continuation are like the continuations in \ineffCEK, but there is no 
constructor for casts.
Continuations in \ineffCEKD\ have zero or more casts at the top.
In \effCEK{C}, however, every continuation has exactly one cast at the top.
Continuations in \ineffCEK\ that have no casts at the top correspond to 
continuations in \effCEK{C}\ whose casts are identity.
And continuations in \ineffCEK\ that have many casts at the top correspond to 
continuations in \effCEK{C}\ where those casts are composed by $seq(c,c)$.

@@

\subsection{Relating Some \effCEK{C} to \ineffCEKD}
\label{secc:framework:monoid-correct}

In this subsection, we prove that if a instance of Cast ADT $C$ is Lazy D and 
is a monoid, then $eval_\mathcal{D} = eval_{\mathcal{S}(C)}$. We proof this 
theorem with a weak bisimulation between \effCEK{C} and \ineffCEKD.

@@

\subsection{\lazyD\ Cast ADT Respect \ineffCEKD}
\label{secc:framework:all-correct}

>>>>

\newcommand{\typingHC}[3]{#1 : #2 \Longrightarrow #3}
\newcommand{\hcvOOinj}[2]{\mathtt{inj} \; #2}
%\newcommand{\hcvOOfun}[5]{\mathtt{fun} \; #1 \; #2 \; #3 \; #4 \; #5}
\newcommand{\hcvOOfun}[5]{\mathtt{fun} \; #2 \; #1 \; #3 \; #4 \; #5}
\newcommand{\hcvOOtt}[0]{\mathtt{unit}}
%\newcommand{\hcvOOcons}[4]{\mathtt{cons}\;#1\;#2\;#3\;#4}
\newcommand{\hcvOOcons}[4]{\mathtt{cons}\;#1\langle#2\rangle\;#3\langle#4\rangle}
%\newcommand{\hcvOOinl}[2]{\mathtt{inl}\;#1\;#2}
%\newcommand{\hcvOOinr}[2]{\mathtt{inr}\;#1\;#2}
\newcommand{\hcvOOinl}[2]{\mathtt{inl}\;#1\langle#2\rangle}
\newcommand{\hcvOOinr}[2]{\mathtt{inr}\;#1\langle#2\rangle}
\newcommand{\hckOOmt}[0]{\mathtt{stop}}
\newcommand{\hckOOconsI}[3]{\mathtt{cons_1}\;#1\;#2\;#3}
\newcommand{\hckOOappII}[2]{\mathtt{app_2}\;#1\;#2}
\newcommand{\sidecond}[1]{\text{if}\;#1}
% Lazy D cast calculus on space-inefficient CEK
\newcommand{\judgeCreduce}[2]{#1 \longmapsto_{\mathcal{C}} #2}
\newcommand{\judgeCreduceTrans}[2]{#1 \longmapsto_{\mathcal{C}}^{*} #2}
\newcommand{\judgeCeval}[2]{eval_{\mathcal{C}}(#1) = #2}
\newcommand{\redrule}[3]{#1 & \longmapsto_\mathcal{C} & #2 & #3\\}
\newcommand{\hiredrule}[3]{\highlight{#1} & \highlight{\longmapsto_\mathcal{C}} & \highlight{#2} & \highlight{#3} \\}
% blame calculus on space-efficient CEK
\newcommand{\judgeSreduce}[3]{#2 \longmapsto_{\mathcal{S}(#1)} #3}
\newcommand{\judgeSreduceTrans}[3]{#2 \longmapsto_{\mathcal{S}(#1)}^{*} #3}
\newcommand{\judgeSeval}[3]{eval_{\mathcal{S}(#1)}(#2) = #3}
\newcommand{\redruleS}[3]{#1 & \longmapsto_{\mathcal{S}(C)} & #2 & #3\\}
\newcommand{\hiredruleS}[3]{\highlight{#1} & 
\highlight{\longmapsto_{\mathcal{S}(C)}} & \highlight{#2} & \highlight{#3} \\}
% Normal Coercion
\newcommand{\ncProj}[2]{#1?^{#2}}
\newcommand{\ncInj}[1]{#1!}
\newcommand{\ncId}[0]{\iota}
\newcommand{\ncSeq}[2]{#1;#2}
\newcommand{\ncFail}[1]{\bot^{#1}}
\newcommand{\ncFun}[2]{\POOfun{#1}{#2}}
\newcommand{\ncProd}[2]{\POOprod{#1}{#2}}
\newcommand{\ncSum}[2]{\POOsum{#1}{#2}}
% Hypercoercion
\newcommand{\hyperCoercionI}[0]{\mathtt{id\star}}
\newcommand{\hyperCoercionC}[3]{#1 \overset{#2}{\curvearrowright} #3}
% machine state simulations
\newcommand{\eqvS}[4]{#3 \approx_{\mathcal{S}\mathcal{S}} #4}
\newcommand{\eqvSD}[3]{#2 \approx_{\mathcal{SD}} #3}
% to-dos
\newcommand{\todoKC}[1]{\todo[inline]{KC needs to #1}}
\newcommand{\todoKCFixed}[0]{\todo[inline]{Fixed. -KC}}
% abbrev
\newcommand{\castCalculus}[0]{$\lambda_{\rightarrow}^{\langle\cdot\rangle}$}
% names
\newcommand{\ineffCEK}{$\mathcal{C}$}
\newcommand{\ineffCEKD}{$\mathcal{D}$}
\newcommand{\ineffCEKUD}{$\mathcal{UD}$}
\newcommand{\judgeDreduce}[2]{#1 \longmapsto_{\mathcal{D}} #2}
\newcommand{\judgeDreduceTrans}[2]{#1 \longmapsto_{\mathcal{D}}^{*} #2}
\newcommand{\judgeDeval}[2]{eval_{\mathcal{D}}(#1) = #2}
\newcommand{\judgeUDreduce}[2]{#1 \longmapsto_{\mathcal{UD}} #2}
\newcommand{\judgeUDreduceTrans}[2]{#1 \longmapsto_{\mathcal{UD}}^{*} #2}
\newcommand{\judgeUDeval}[2]{eval_{\mathcal{UD}}(#1) = #2}
\newcommand{\effCEK}[1]{$\mathcal{S}(#1)$}
\newcommand{\evalEqv}[2]{\ensuremath{eval_{\text{#1}} = eval_{\text{#2}}}}
\newcommand{\continue}[2]{cont(#2,#1)}
\newcommand{\highlight}[1]{{\color{red} #1}}

@@

\begin{definition}[\lazyD{} CEK Machine]
  The \lazyD{} CEK machine, written \ineffCEKD{}, is the CEK machine of
  \figref{fig:machine-cekc} using the $applyCast$ for \lazyD{} defined
  in \figref{fig:applyCast-D-C}.  We write the transition relations of
  this machine as $\judgeDreduce{s}{s}$ and $\judgeDreduceTrans{s}{s}$
  and write the evaluation function as $\judgeDeval{e}{o}$.
\end{definition}

@@

\begin{definition}[\lazyUD{} CEK Machine]
  The \lazyUD{} CEK machine, written \ineffCEKUD{}, is the CEK machine of
  \figref{fig:machine-cekc} using the $applyCast$ for \lazyUD{} defined
  in \figref{fig:apply-Cast-UD}.  We write the transition relations of
  this machine as $\judgeUDreduce{s}{s}$ and $\judgeUDreduceTrans{s}{s}$
  and write the evaluation function as $\judgeUDeval{e}{o}$.
\end{definition}

@@

This section presents our second contribution, a framework for proving
the correctness of cast representations, especially space efficient
ones. Section~\ref{sec:hypercoercion-correctness} applies this
framework to prove the correctness of the \lazyD{} hypercoercions.

We start by defining a CEK machine \effCEK{C} that is parameterized
over the cast representation $C$
(Section~\ref{sec:framework:cek}). This machine is space efficient
provided that the cast representation performs compression.  The
interface between the machine and the cast representation is defined
by an abstract data type, named Cast ADT, that we define shortly.  We
conjecture that all cast representations defined in the literature
are instances of this ADT.

In Section~\ref{secc:framework:monoid-correct} we prove that, for any
instance $C$ of the Cast ADT, if $C$ satisfies a more refined abstract
data type for Lazy D casts, then \effCEK{C} is equivelent to
\ineffCEKD{}, that is,
\[
  \evalEqv{\effCEK{C}}{\ineffCEKD}
\]
We conjecture that all of the cast representations in the literature
for \lazyD{} (supercoercions, coercions in normal form, threesomes)
are instances of the Lazy D Cast ADT.  We are working on a similar
theorem for \lazyUD{} cast representations.

We then apply this framework to \lazyD{} hypercoercions
(Section~\ref{sec:hypercoercion-correctness}), where we show that $H$
is an instance of the Lazy D Cast ADT, and therefore
\[
  \evalEqv{\effCEK{H}}{\ineffCEKD}
\]

Before turning to the definition of the abstract data types, we first
give the definition of values used by the \effCEK{C}, as they are
mentioned in the definition of the ADTs.

\begin{definition}[Values for the \effCEK{C} machine] \label{def:values-effCEK}
  \[
  \begin{array}{rclr}
	\stxrule{v}{}{
		\hcvOOtt \mid
		\hcvOOfun{c}{\rho}{x}{e}{c} \mid
		\hcvOOcons{v}{c}{v}{c} \mid
		\hcvOOinl{v}{c} \mid
		\hcvOOinr{v}{c} \mid
		\hcvOOinj{P}{v}
	}
  \end{array}
  \]
\end{definition}

\todo[inline]{We also need to define results $r$ for \effCEK{C} -JS}

\subsection{The Cast Abstract Data Types}

The \emph{Cast Abstract Data Type}, defined below, captures the set of
operators that a cast representation $C$ must provide for it to be
used with the \effCEK{C} machine.  The first three operators enables
\effCEK{C} to construct casts so we call them \textit{cast
  constructors}. The forth and last operator enables \effCEK{C} to
apply casts to values.

\begin{definition}[Cast Abstract Data Type (Cast ADT)]
  \label{def:cast-rep}
  A cast abstract data type is a set $Cast$, which is indexed by two types, 
  with four operators:
  \begin{description}
  \item[$id(T)$] constructs an identity cast from a type
  \item[$seq(c_1,c_2)$] composes two casts
  \item[$cast(T_1,l,T_2)$] constructs a cast from $T_1$ to $T_2$
  \item[$ applyCast(v,c)$] applies a cast to a value, producing a result
  \end{description}
  We denote by $c : T_1 \Longrightarrow T_2$ to mean $c$ is in $Cast \; T_1 \; T_2$
  and we say ``$c$ is from $T_1$ to $T_2$''.
\end{definition}

Next we define the \emph{Lazy D Cast ADT}, which capture the further
requirements that are needed for the theorem that establishes
equivalence to \ineffCEKD{}. Property (1) below states that $id(T)$
acts like the identity function. Property (2) states that $seq(c,c)$
acts like a sequence of casts. Properties (3) through (11) state that
$cast(T_1,l,T_2)$ act like the Lazy D $applyCast$
(Fig.~\ref{fig:applyCast-D-C}).

\begin{definition}[Lazy D Cast ADT]
  \label{def:surely-lazyd}
  A  Cast is a Lazy D Cast if:
  \begin{enumerate}
  \item if $v : T$, then $applyCast(v,id(T)) = \mathtt{succ} \; v $
  \item if $\judgeType{v}{T_1}$,
    $ \judgeTypeFT{c_1}{T_1}{T_2}$, and
    $ \judgeTypeFT{c_2}{T_2}{T_3}$,\\
    then $applyCast(v,seq(c_1,c_2)) = 
    applyCast(v,c_1) >>= \lambda v.applyCast(v,c_2)$\\
    where 
    \[
    \begin{array}{rcl}
      \rOOsucc{v} >>= f & = & f(v) \\
      \rOOfail{l} >>= f & = & \rOOfail{l}
    \end{array}
    \]
  \item if $v : T_1$ and $\neg T_1 \smile T_2$,
    then $applyCast(v,cast(T_1, l, T_2)) = \rOOfail{l} $
  \item if $v : \star$, 
    then $applyCast(v,cast(\TOOdyn,l,\TOOdyn)) = \rOOsucc{v} $
  \item if $v : P$,
    then $applyCast(\hcvOOinj{P}{v},cast(\star,l,P')) 
    = applyCast(v,cast(P,l,P')) $
  \item if $v : P$,
    then $applyCast(v,cast(P,l,\star)) = \rOOsucc{(\hcvOOinj{P}{v})} $
  \item if $v : \POOunit$,
    then $applyCast(v,cast(\POOunit,l,\POOunit)) = \rOOsucc{v} $
  \item if $(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}) : \POOfun{T_1}{T_2}$,
    then\\
    $ 
    applyCast(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}, 
    cast(\POOfun{T_1}{T_2},l,\POOfun{T_3}{T_4})) \\
    = 
    \rOOsucc{(\hcvOOfun{seq(cast(T_3,l,T_1),c_1)}{\rho}{x}{b}{seq(c_2,cast(T_2,l,T_4))})}$
  \item if $(\hcvOOcons{v_1}{c_1}{v_2}{c_2}) : \POOprod{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOcons{v_1}{c_1}{v_2}{c_2},cast(\POOprod{T_1}{T_2},l,T_3 
    \times 
    T_4))$\\
    $ = 
    \rOOsucc{(\hcvOOcons{v_1}{seq(c_1,cast(T_1,l,T_3))}{v_2}{seq(c_2,cast(T_2,l,T_4))})}
    $ 
  \item if $(\hcvOOinl{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOinl{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))})} $
  \item if $(\hcvOOinr{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\$
    applyCast(\hcvOOinr{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinr{v}{seq(c,cast(T_2,l,T_4))})} $
  \end{enumerate}
\end{definition}

%% After showing their cast is an instance of Lazy D Cast, users of our
%% frameworks can plug-in this cast representation to our space-efficient
%% abstract machine in subsection~\ref{ssec:framework:cek}, and prove
%% that this machine is equivalent to
%% \ineffCEKD\ (subsection~\ref{secc:framework:all-correct}), which
%% depends on lemmas in subsection~\ref{secc:framework:monoid-correct}.

@@

\begin{figure}
	Syntax
	\[
	\begin{array}{rclr}
%	\stxrule{r}{cast results}{
%		\rOOsucc{v} \mid
%		\rOOfail{l}
%	}
%	\stxrule{s}{states}{
%		\sOOinspect{e}{\rho}{\kappa} \mid{}
%		\sOOreturn{v}{\kappa} \mid{}
%		\sOOhalt{o}
%	}
	\stxrule{\kappa}{continuation}{
		\kOOcast{c}{k}
	}
	\stxrule{k}{pre-continuations}{
		\kOOmt \mid{}
		\kOOconsI{e}{\rho}{\kappa} \mid
		\kOOconsII{v}{\kappa} \mid
		\kOOinl{\kappa} \mid
		\kOOinr{\kappa}
	}
	\stxrulecont{
		\kOOappI{e}{\rho}{\kappa} \mid
		\kOOappII{v}{\kappa} \mid
		\kOOcar{\kappa} \mid
		\kOOcdr{\kappa}
	}
	\stxrulecont{
		\kOOcaseI{e}{e}{\rho}{\kappa} \mid
		\kOOcaseII{v}{e}{\rho}{\kappa} \mid
		\kOOcaseIII{v}{v}{\kappa}
	}
	\end{array}
	\]

@@

	\caption{Space-efficient CEK machine$\mathcal{S}(C)$}
	\label{fig:machine-cekcc}
\end{figure}

\figref{fig:machine-cekcc} defines \effCEK{C}, a space-efficient CEK
machine.  This machine is similar to the \ineffCEK{} machine
(Figure~\ref{fig:machine-cekc}).  The key differences are in values
and continuations.

Let $v$ range over values. In \ineffCEK{} we have one value
constructor for all casted value.  In \effCEK{C}, however, we do not
have this generic value constructor, instead we push those casts into
the ordinary values. For instance, $(\vOOcons{v_1}{v_2})$ corresponds
to $(\hcvOOcons{v_1'}{id(T_1)}{v_2'}{id(T_2)})$. To construct values
of the dynamic type, we use $(\hcvOOinj{P}{v})$.

Cast results ($r$) and machine states ($s$) are the same as for
\ineffCEK{}.

Let $\kappa$ range over continuations and let $k$ range over
pre-continuations.  A continuation is now a pre-continuation prefixed
with a cast.  Pre-continuation are like the continuations in
\ineffCEK{}, but there is no constructor for casts.  Continuations in
\ineffCEKD{} have zero or more casts at the top.  In \effCEK{C},
however, every continuation has exactly one cast at the top.
Continuations in \ineffCEK\ that have no casts at the top correspond
to continuations in \effCEK{C}\ whose casts are identity.
Continuations in \ineffCEK\ that have many casts at the top correspond
to continuations in \effCEK{C}\ where those casts are composed by
$seq(c,c)$.

@@

\subsection{\effCEK{C} is equivalent to \ineffCEKD, provided $C$ is an instance of the Lazy D ADT}
\label{secc:framework:monoid-correct}

In this subsection, we prove that for all $C$, if $C$ is a Lazy D Cast
ADT, then
\[
  \evalEqv{\effCEK{C}}{\ineffCEKD}
\]

The meat of the proof is in a lemma that establishes a weak
bisimulation between \ineffCEKD{} and \effCEK{C}.

We prove that if a instance of Cast ADT $C$ is
Lazy D and is a monoid, then $eval_\mathcal{D} =
eval_{\mathcal{S}(C)}$. We proof this theorem with a weak bisimulation
between \effCEK{C} and \ineffCEKD.

@@

%% \subsection{\lazyD\ Cast ADT Respect \ineffCEKD}
%% \label{secc:framework:all-correct}
